---
title: "Query Pubmed by Grant"
author: "Nicole Kauer", "Kelsey Montgomery"
date: "3/24/2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dccvalidator)
library(dplyr)
library(easyPubMed)
library(readr)
library(reticulate)
library(porTools)
synapseclient <- reticulate::import("synapseclient")
syntab <- reticulate::import("synapseclient.table")
syn <- synapseclient$Synapse()
syn$login()
```

## Query Pubmed and store data as file annotations

The data needed for these steps are *Grant Number*, *grantSerialNumber* and *Program*. Theses functions take a list of grant serial numbers and queries Pubmed to download title, abstract, authors, journal name, year and DOI. Theses annotations are visible in the [AD Knowledge Portal - Publications View](https://www.synapse.org/#!Synapse:syn20448807/tables/). See the [Explore Publications module](https://adknowledgeportal.synapse.org/Explore/Publications) for a visual of how this data is surfaced on the portal.

```{r ex_format, echo = FALSE}
tribble(~`Grant Number`, ~grantSerialNumber, ~Program,
        "U01AG046139", "AG046139", "AMP-AD"
        )
```

Import the grants with their respective programs and serial numbers.

```{r vars, echo=FALSE}
table_id <- "syn17024229"
grants <- syn$tableQuery(
  glue::glue("SELECT \"Grant Number\", grantSerialNumber, Program FROM {table_id}")
)$asDataFrame()
# Remove rows that have NaN or NA or empty string for the serial number
grants <- grants[!(grants$grantSerialNumber %in% c(NaN, NA, "")), ]
```
## Run the code

Any character vector can be passed to `query_list_general`. This function wraps several functions:
- query Pubmed
- create an entity name from first author, journal, year and Id
- abbreviates the author names by first initial, last name
- creates one row per PubmedId
- creates grant column to associate with PubmedId
- leaves out grants that were not associated with a PubmedId
- creates a *query* column to associate the PubmedId with a specific query

```{r query, message=FALSE, warning=FALSE}
dat <- query_list_general(grants$grantSerialNumber)
```

Join the grants to the Pubmed queries and clean up.

```{r query}
dat <- dat %>%
  rename(grantSerialNumber = query)
# For some reason, grantSerialNumber isn't always a character
grants$grantSerialNumber <- as.character(grants$grantSerialNumber)
dat <- dplyr::right_join(grants, dat, by = "grantSerialNumber")
# Need to remove duplicates, but keep all grants and consortium
# Includes some renaming and dropping of columns
dat <- dat %>%
  group_by(pmid) %>%
  mutate(grant = glue::glue_collapse(unique(.data$`Grant Number`), ", ")) %>%
  mutate(consortium = glue::glue_collapse(unique(.data$Program), ", ")) %>%
  select(!c(`Grant Number`, Program, grantSerialNumber)) %>%
  rename(pubmed_id = pmid, DOI = doi, Program = consortium) %>%
  distinct()
```

The following has fixes for some of the formatting issues found. It also updates the entity name to remove common, unallowed characters.

```{r hacky}
# Included in hacky_cleaning is conversion to ascii and removing html formatting
dat$title <- hacky_cleaning(dat$title)
dat$authors <- hacky_cleaning(dat$authors)
dat$journal <- hacky_cleaning(dat$journal)
dat$abstract <- hacky_cleaning(dat$abstract)
# Remove common, unallowed characters from entity name; includes hacky_cleaning
dat$entity_name <- remove_unacceptable_characters(dat$entity_name)
```

`set_up_multiannotations` parses comma-separated lists to be stored correctly in Synapse as multi-value annotations. Before setting up the multiannotations, add extra columns that are needed for working with the Portal. The additional, redundant columns will be removed in the future. Should keep `grant` and `Program`, and remove `long_amp_ad_grants`, `doi`, and `consortium`.

```{r columns}
dat <- set_up_multiannotations(dat, "grant")
dat <- set_up_multiannotations(dat, "Program")
```

The final data is transposed so that it can be iterated over by `purrr` and stored in Synapse under the `parent` folder.

```{r store, message=FALSE, echo=FALSE}
parent = "syn20463015"
dat_list <- purrr::transpose(dat)
store_as_annotations(parent = parent, dat_list)
```

Query the publications table to force an update.

```{r query, message=FALSE, echo=FALSE}
pub_table <- "syn20448807"
syn$tableQuery(glue::glue("SELECT * FROM {pub_table} LIMIT 1"))
```
