---
title: "Query Pubmed"
author: "Kelsey Montgomery, Nicholas Lee"
date: "9/18/2020"
edited: "04/28/2023"
notes: "Adapted for the ELITE portal"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(dccvalidator)
library(dplyr)
library(easyPubMed)
library(readr)
library(reticulate)
library(synapser)

synLogin()
```

## Query Pubmed and store data as file annotations

The schema for these steps is *pubmedId*, *grants* and *study*. Theses functions take
a list of PubmedIds and queries the site to pull down title, abstract, authors, journal name, year and DOI. Theses annotations are visible in the [ELITE Portal - Publications View](https://www.synapse.org/#!Synapse:syn51209321/tables/). See the [Explore Publications module](https://psychencode.synapse.org/Explore/Publications) for a visual of how this data is surfaced on the portal.

```{r query synapse for grants}
sid_elite <- "syn27229419"
sid_studies <- "syn51209885"

grant_list <-
  readr::read_csv(syn$tableQuery(paste(
    "SELECT * FROM ", sid_studies, sep = ""
  ))$filepath,
  col_types = readr::cols(.default = "c"))

# cleanup grant list
grant_list$grantNumber <- gsub('\\[|\\]|\\"',"", as.character(grant_list$grantNumber))

# take only grant number as the grants
grants <- grant_list$grantNumber
```


```{r ex_format, echo = FALSE}
# taking pubmedIDs from file within synapse

# tribble(~pubmedId, ~grants, ~study,
#         "24057217", "R21MH103877", c("study1,study2")
#         )
```

Import the list of Pubmed Ids and define the Synapse parentId where the file entities will be stored with the Pubmed-relevant annotations.
```{r vars, echo=FALSE}
parent <- "syn51400816" # ELITE Portal backend folder for publications

# read pmids file
pmids <- readr::read_tsv(syn$get("syn22080024")$path, 
                         col_types = readr::cols(.default = "c"))
```

## Run the code

Any character vector can be passed to `query_list_pmids`. This function wraps several functions:
- query Pubmed
- create an entity name from first author, journal, year and Id
- abbreviates the author names by first initial, last name
- creates one row per PubmedId

```{r query}
dat <- query_list_pmids(pmid_map$pubmed_Id)
```

I am keeping an eye out for weird edge cases. These (hacky) steps clean some missing values and remove extraneous characters.

```{r hacky}
dat$title <- gsub("<i>|</i>", "", dat$title)
dat$authors <- gsub("NA", "", dat$authors)
dat$entity_name <- gsub("NA ", "", dat$entity_name)
dat$journal <- remove_unacceptable_characters(dat$journal)
dat$entity_name <- remove_unacceptable_characters(dat$entity_name)
```

`set_up_multiannotations` parses comma-separated lists to be stored correctly in Synapse as multi-value annotations. Then, study and grant annotations are joined to the Pubmed queries.

```{r join}
pmids <- set_up_multiannotations(pmids, "study")
pmids <- set_up_multiannotations(pmids, "grants")

mapping <- dplyr::right_join(pmids, dat, by = c("pubmedId" = "pmid"))
mapping <- dplyr::rename(mapping, DOI = doi)
```

The final data is transposed so that it can be iterated over by `purrr` and stored in Synapse.

```{r store}
list <- purrr::transpose(mapping)
store_as_annotations(parent = "syn22235314", list)
```
